---
title: "Beaujean Modelling in R for SEM"
author: "Alex Brooks"
date: "10/15/2018"
output: html_document
---

```{r setup, include=FALSE}
library(lavaan)
library(semPlot)
options (scipen = 999)
library(dplyr)
library(caret)
library(corrplot)
library(glmnet)
library(gapminder)
library(openintro)
library(ggplot2)
library(fields)
library(dagitty)
```
 
##First try making a DAG  - a graphic to explain my hypothesis

x = pc_immun_code (percent_immunised sa3, pc immun rank by postcode)
y = geo (state, postcode, sa3, electorate - could do clusters)
z = socio (SEIFA score)
p = political  - political score
t = tax - mean tax 000s
```{r}
g <- dagitty('dag {
    X [pos="0,1"]
    Y [pos="1,1"]
    Z [pos="2,1"]
    P[pos="1,0"]
    T [pos="2,2"]
    
    X -> Y -> Z  
    X -> Z -> Y -> T -> P
    P -> Z
    X -> P -> Z -> Y
}')
plot(g)
```

```{r}
#name all the parents of x
parents( g, "P" )
#these are T and X
```
```{r}
#name all the ancestors of x
ancestors( g, "P" )
#"P" "X" "T" "Y" "Z"
```
```{r}
#name all the childrens of P
children( g, "P" )
#z
```
```{r}
#name all the descendents of P
descendants( g, "P" )
# "P" "Z" "Y" "T"
```
```{r}
#draw all simple paths between P and T
paths( g, "P", "T" )$paths

```
```{r}
#draw all the directed paths between P and T
paths( g, "P", "T", directed=TRUE )$paths
```
### Dagitty study questions
x = pc_immun_code (percent_immunised sa3, pc immun rank by postcode)
y = geo (state, postcode, sa3, electorate - could do clusters)
z = socio (SEIFA score)
p = political  - political score
t = tax - mean tax 000s

Does tax, political, socio or geo cause x? 
```{r}
N <- 2484 # 2484 is our sample size, as that's the number of postcodes and entries
Ux <- rnorm( N ); Uy <- rnorm( N ); Uz <- rnorm( N )
X <- Ux
Y <- 1/3*X + Uy
Z <- 1/16*Y + Uz
d <- data.frame(X=X,Y=Y,Z=Z)
```

```{r}
g <- dagitty("dag {
    Ux -> X -> Y -> Z <- Uz
    Uy -> Y
}")
coordinates(g) <- list(
    x=c(Ux=1,Uy=2,Uz=3,X=1,Y=2,Z=3),
    y=c(Ux=1,Uy=1,Uz=1,X=0,Y=0,Z=0) )
plot(g)
```
```{r}
#determine the best guess of the value of z if we observe y = 3
16 * predict(lm(Z~Y,d),list(Y=3),interval="confidence")
```
```{r}
#Determine the best guess of the value of Z, given that we observe X=1 and Y=3.
16 * predict(lm(Z~X+Y,d),list(X=1,Y=3),interval="confidence")
```
```{r}
#Determine the best guess of X, given that we observed Y=2.
cov(d)

```
#The regression coefficient of X∼Y should be 13/(1+19)=930=0.3. So the best guess should be 0.6.
```{r}
predict(lm(X~Y,d),list(Y=2),interval="confidence")
```
#Determine the best guess of Y, given that we observed X=1 and Z=3.
```{r}
#Finding this value requires an involved calculation, which gives an answer close to (but not equal to) 0.5. In case you wish to verify your result, here’s how:
predict(lm(Y~X+Z,d),list(X=1,Z=3),interval="confidence")
```


## First, bring in the 2016 postcode-level data, and analyse the 5 year olds only

Start with 2016 only data, which has had funding per capita added, political score added and reduced SEIFA scores created.

```{r 2016}
 #read in the data for 2016 only
Data <-read.csv('../cleaned_data/2016_data_reduced.csv')
str(Data)
View(Data)
```


```{r per capita funding workings 2016}
#calculate government funding per postcode
#Budget papers say the Feds spent $6.4m in 2016 on improving childhood immunisation rates
#Total population of Australia in 2016 was 24 598 900 according to the ABS
#We have the SEIFA URP which has the usual resident population per postcode
#First get per person amount
total_population <- (24598900)
total_immun_spend <- (6400000)
per_capita <- total_population/total_immun_spend
per_capita
#3.843578 spend per person

```
```{r cleaning}
# Remove NA values
Data <- Data[complete.cases(Data), ]

```
#How many postcodes are there in the 2016 set
2484 postcodes in 2016
150 electorates in 2016
31 PHNs in 2016
12 states in 2016
```{r}
#count postcodes
length(unique(Data$postcode))
#count eletorates
length(unique(Data$electorate))
#count PHNs
length(unique(Data$PHN_number))
#count states - because there are doubles from the weird AIHW counting
length(unique(Data$state))
```

```{r}
#Need to make the factors ordered factors to work in Lavaan
Data$postcode = postcode_factor_ordered <- factor(Data$postcode, ordered = TRUE)
Data$state = state_factor_ordered <- factor(Data$state, ordered = TRUE)
Data$pc_immun = pc_immun_factor_ordered <- factor(Data$pc_immun, ordered = TRUE)
Data$PHN_code = PHN_code_factor_ordered <- factor(Data$PHN_code, ordered = TRUE)
Data$electorate = electorate_factor_ordered <- factor(Data$electorate, ordered = TRUE)
Data$govt_funds_per_person = funds_ordered <- as.numeric(Data$govt_funds_per_person)
glimpse(Data)
```

```{r, fig.width=18, fig.height=12}
 
#correlation plot - this won't work, I think due to NAs - this works in OCt15 modelling doc
cplot <- Data %>%
  select_if(is.numeric)


M <- cor(cplot)
mat <- cor.mtest(M)
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
corrplot(M, 
         method = "color",
         order= "hclust",
         type="full",
         col=col(200),
         diag =F,
         title="Correlation of Numeric Variables",
         addCoef.col = "black",
         sig.level = 0.05,
         insig ="blank",
         mar=c(0,0,3,0))

```


```{r}
#filter just by 5 year olds - as the per capita info in spreadsheet is duplicating by age group
Five_Data = Data%>%
  filter(age == 5) 

  
str(Five_Data)
```
```{r}
#let's look at the variance of this dataset
varTable(pc.fit)
```


#First 2016 model - assessing a real value against the other real values (IE, not latent yet)

  # single factor model
    geo =~ pc_immun_class + IRSAD_SCORE + political_score + govt_funds_per_person
    

```{r}

pc.model <- 'geo =~ pc_immun_class + IRSAD_SCORE + political_score + govt_funds_per_person'
pc.fit <- cfa(model=pc.model,
                  data = Five_Data, std.lv = TRUE)
summary(pc.fit, standardized = TRUE, fit.measures = TRUE)

```
```{r}
lavInspect(pc.fit, what="estimates")
```

```{r}
semPaths(pc.fit,  what = "stand", rotation = 4)
```

You can also embed plots, for example:

```{r}
semCors(pc.fit)
```

#Try the same single factor model as first model usng 2015 data  
```{r 2015}
 #read in the data for 2015 only
#try this to stop data confusion
char_data <- read.csv('../cleaned_data/2015_dataREDUCED.csv')
str(char_data)
```
```{r}
#Need to make the factors ordered factors to work in Lavaan
char_data$postcode = postcode_factor_ordered <- factor(char_data$postcode, ordered = TRUE)
char_data$state = state_factor_ordered <- factor(char_data$state, ordered = TRUE)
char_data$pc_immun = pc_immun_factor_ordered <- factor(char_data$pc_immun, ordered = TRUE)
char_data$PHN_code = PHN_code_factor_ordered <- factor(char_data$PHN_code, ordered = TRUE)
char_data$electorate = electorate_factor_ordered <- factor(char_data$electorate, ordered = TRUE)
char_data$per_capita = funds_ordered <- as.numeric(char_data$per_capita)
glimpse(char_data)
```

 
 Calculate per capita spend 
 Govt spent $8.4m, population total was 24 127 200 in 2015-16

#calculate government funding per postcode
 #We have the SEIFA URP which has the usual resident population per postcode
#First get per person amount
```{r}
total_population15 <- (24127200)
total_immun_spend15 <- (8400000)
per_capita15 <- total_population15/total_immun_spend15
per_capita15
#2.87228 spend per person
```

Filter for 5 year olds
```{r}
#filter just by 5 year olds - as the per capita info in spreadsheet is duplicating by age group
Five_Data15 = char_data%>%
  filter(age == 5) 

  
str(Five_Data15)
```

```{r}
#Our geo factor is measured by PHN, IRSAD score and pc immun class
geo_2015.model <- 'socio =~ mean_tax_000s + political_score + IRSAD_SCORE + per_capita + pc_immun_class'
geo_2015.fit <- cfa(model = geo_2015.model,
                  data = Five_Data15)
summary(geo_2015.fit, standardized = TRUE, fit.measures = TRUE)
  
```


```{r}
lavInspect(geo_2015.fit, what="estimates")
```
```{r}
semPaths(geo_2015.fit,  what = "stand", rotation = 4)
```

You can also embed plots, for example:

```{r}
semCors(geo_2015.fit)
```
```{r}
#modification indices (rmsea and srmr need to be below 0.10)
#look at loading and variances to check whether they are improbable or just don't relate to the latent variable, if loadings are over point 3 (variance will be large in comparison to the data)
#var(final_data)

```

# Let's try SA3 data and see if it can get a model working
```{r}
sa3 <-read.csv('../cleaned_data/sa3_immuneMERGED.csv')
str(sa3)
```


```{r cleaning}
# Remove NA values
sa3 <- sa3[complete.cases(sa3), ]

```


```{r}
#Need to make the factors ordered factors to work in Lavaan
sa3$state = state_factor_ordered <- factor(sa3$state, ordered = TRUE)
sa3$num_registered = num_registered_factor_ordered <- as.numeric(sa3$num_registered, ordered = TRUE)
sa3$num_fully_immunised = imm_factor_ordered <- as.numeric(sa3$num_fully_immunised, ordered = TRUE)
sa3$num_not_fully_immunised = not_imm_factor_ordered <- as.numeric(sa3$num_not_fully_immunised, ordered = TRUE)
sa3$percent_fully_immunised = percent_ordered <- factor(sa3$percent_fully_immunised, ordered = TRUE)
sa3$Remoteness = Remoteness_ordered <-factor(sa3$Remoteness, ordered = TRUE)
glimpse(sa3)
 
```
```{r}
#filter for 5 year olds like the other dataset
Five_sa3 = sa3%>%
  filter(age == 5) %>%
  filter(pc_immun)

  
str(Five_sa3)

```


#Try building a model with SA3 level data
```{r}
#sa3
sa3.model <- 'score =~ percent_fully_immunised + Remoteness+'
sa3.fit <- cfa(model=sa3.model,
                  data = Five_sa3)
summary(sa3.fit, standardized = TRUE, fit.measures = TRUE)
  
```
```{r}

semCors(trust.fit)
```

```{r}
semPaths(trust.fit,  what = "stand", rotation = 4)
```
```{r}
semCors(trust.fit, include="difference")
```
##This isn't really working
Let's see if we can replicate the geo modelling at SA3 level

```{r}
 
#Try a single factor model
#Our geo factor is measured by PHN, IRSAD score and pc immun class
sa3.model <- 'sa3 =~ SA3_code + year_start + age + num_not_fully_immunised + num_fully_immunised + percent_fully_immunised'
sa3.fit <- cfa(model=sa3.model,
                  data = final_data)
summary(sa3.fit, standardized = TRUE, fit.measures = TRUE)
  
```

```{r}
#plot it
lavInspect(sa3.fit, what="estimates")
```
```{r}
semPaths(sa3.fit,  what = "stand", rotation = 4)
```
```{r}
semCors(sa3.fit, include="difference")
```
