---
title: "SEM immunisation data in Australia, 2015-16 and 2016-17"
author: "Alex Brooks"
date: "10/15/2018"
output: html_document
---

```{r setup, include=FALSE}
library(lavaan)
library(semPlot)
options (scipen = 999)
library(dplyr)
library(corrplot)
library(ggplot2)
library(fields)
library(dagitty)
```
 
##First explain the model using 2016-17 immunisation data at postcode level, then build 2015-16 immunisation data at postcode level
We are trying to answer the question: Do geographic location, socio-economics, political affiliation or government funding have a relationship with immunisation coverage?

We will do this by using SEM on two years of postcode-level immunisation data to either confirm or deny whether a relationship exists.

Latent variables
% immunisation - this is confusing because this is a manifest variable when used as pc_immun_class , but also a constructed latent variable as it's the 'target' of interest. 
x =immunisation coverage (as measured by manifest variables of PC immun code   
y = geo (Geographic areas or clusters, as measured by state, postcode, sa3, electorate)
z = socio-economics (as measured by SEIFA score and mean tax)
p = political  (as measured by political score)
f = funding (as measured by immunisation per person in that postcode or sa3)
```{r 2016}
 #read in the data for 2016 only
Data <-read.csv('../cleaned_data/2016_data_REDUCED.csv')
str(Data)
View(Data)
```


```{r per capita funding workings 2016}
#calculate government funding per postcode
#Budget papers say the Feds spent $6.4m in 2016 on improving childhood immunisation rates
#Total population of Australia in 2016 was 24 598 900 according to the ABS
#We have the SEIFA URP which has the usual resident population per postcode
#First get per person amount
total_population <- (24598900)
total_immun_spend <- (6400000)
per_capita <- total_population/total_immun_spend
per_capita
#3.843578 spend per person

```
```{r cleaning}
# Remove NA values
Data <- Data[complete.cases(Data), ]

```
#How many postcodes are there in the 2016 set
2484 postcodes in 2016
150 electorates in 2016
31 PHNs in 2016
12 states in 2016
```{r}
#count postcodes
length(unique(Data$postcode))
#count eletorates
length(unique(Data$electorate))
#count PHNs
length(unique(Data$PHN_number))
#count states - because there are doubles from the weird AIHW counting
length(unique(Data$state))
```

```{r}
#Need to make the factors ordered factors to work in Lavaan
Data$postcode = postcode_factor_ordered <- factor(Data$postcode, ordered = TRUE)
Data$state = state_factor_ordered <- factor(Data$state, ordered = TRUE)
Data$pc_immun = pc_immun_factor_ordered <- factor(Data$pc_immun, ordered = TRUE)
Data$PHN_code = PHN_code_factor_ordered <- factor(Data$PHN_code, ordered = TRUE)
Data$electorate = electorate_factor_ordered <- factor(Data$electorate, ordered = TRUE)
Data$govt_funds_per_person = funds_ordered <- as.numeric(Data$govt_funds_per_person)
glimpse(Data)
```


```{r}
#filter just by 5 year olds - as the per capita info in spreadsheet is duplicating by age group
Five_Data = Data%>%
  filter(age == 5) 

  
str(Five_Data)
```



#First 2016 model - assessing a real value against the other real values (IE, not latent yet)

  # single factor model
    imm =~ pc_immun_class + IRSAD_SCORE + political_score + govt_funds_per_person + mean_tax_000s
    

```{r}
#let's build 2016pc.model the standard way
pc2016.model <- 'imm =~ pc_immun_class + IRSAD_SCORE + political_score + govt_funds_per_person + mean_tax_000s'
pc2016.fit <- cfa(model=pc2016.model,
                  data = Five_Data, std.lv = TRUE)
summary(pc2016.fit, standardized = TRUE, fit.measures = TRUE)
```
```{r}
#let's look at the variance of this dataset
varTable(pc2016.fit)
```

```{r}
#extracts the estimated values of the free parameters in the model
coef(pc2016.fit)
```

```{r}
lavInspect(pc2016.fit, what="estimates")
```

```{r}

semPaths(pc2016.fit, title=TRUE, exoVar = TRUE, exoCov = TRUE)
```
```{r}
semPaths(pc2016.fit,  what = "stand", rotation = 4)
```

```{r}
#
semCors(pc2016.fit, titles = TRUE)
```
```{r}
 #let's build to account for missing pc immun data
pc2016.fit2 <- cfa(model=pc2016.model,
                  data = Five_Data, estimator = "MLR", missing = "ml")
summary(pc2016.fit2, standardized = TRUE, fit.measures = TRUE)
#THIS IS THE ONE TO PUBLISH AS MODEL FIT
```
```{r}
#Let's bootstrap the model
pc2016.fit3 <- cfa(model=pc2016.model,
                  data = Five_Data, se = "bootstrap", verbose = TRUE, bootstrap = 1000)

```
```{r}
#fitmeasures
fitMeasures(pc2016.fit2)
```
```{r}
#fit measures for first simple model)
fitMeasures(pc2016.fit)
```
```{r}
#parameter table
parameterTable(pc2016.fit)
```
##Build the single factor 2015 model
```{r}
 #read in the data for 2015 only
#try this to stop data confusion
char_data <- read.csv('../cleaned_data/2015_dataREDUCED.csv')
str(char_data)
```
```{r}
#Need to make the factors ordered factors to work in Lavaan
char_data$postcode = postcode_factor_ordered <- factor(char_data$postcode, ordered = TRUE)
char_data$state = state_factor_ordered <- factor(char_data$state, ordered = TRUE)
char_data$pc_immun = pc_immun_factor_ordered <- factor(char_data$pc_immun, ordered = TRUE)
char_data$PHN_code = PHN_code_factor_ordered <- factor(char_data$PHN_code, ordered = TRUE)
char_data$electorate = electorate_factor_ordered <- factor(char_data$electorate, ordered = TRUE)
char_data$per_capita = funds_ordered <- as.numeric(char_data$per_capita)
glimpse(char_data)
```

 
 Calculate per capita spend 
 Govt spent $8.4m, population total was 24 127 200 in 2015-16

#calculate government funding per postcode
 #We have the SEIFA URP which has the usual resident population per postcode
#First get per person amount
```{r}
#total_population15 <- (24127200)
#total_immun_spend15 <- (8400000)
#per_capita15 <- total_population15/total_immun_spend15
#per_capita15
#2.87228 spend per person
```

Filter for 5 year olds
```{r}
#filter just by 5 year olds - as the per capita info in spreadsheet is duplicating by age group
Five_Data15 = char_data%>%
  filter(age == 5) 
#make sure all are ordered factors, as data seems to have changed from previous str
Five_Data15$per_capita = funds_ordered <- as.numeric(Five_Data15$per_capita)
glimpse(Five_Data15)
str(Five_Data15)
```

```{r}
#Our geo factor is measured by PHN, IRSAD score and pc immun class
geo_2015.model <- 'imm =~ mean_tax_000s + political_score + IRSAD_SCORE + per_capita + pc_immun_class + per_capita'
geo_2015.fit <- cfa(model = geo_2015.model,
                  data = Five_Data15)
summary(geo_2015.fit, standardized = TRUE, fit.measures = TRUE)
  
```


```{r}
lavInspect(geo_2015.fit, what="estimates")
```
```{r}
semPaths(geo_2015.fit,  what = "stand", rotation = 4)
```

```{r}
semCors(geo_2015.fit)
```


```{r}

semCors(geo_2015.fit, titles = TRUE)
```
```{r}
coef(geo_2015.fit)
```
##Merging 2015 and 2016 data into one variable
##Now let's build a more complete model using the two years to see if things can be confirmed
```{r}

colnames(Five_Data15)[names(Five_Data15) == "per_capita"] <- "govt_funds_per_person"
colnames(Five_Data)[names(Five_Data) == "SEIFA_URP"] <- "IRSAD_URP"

all_data <- rbind(Five_Data15, Five_Data)

all_data$postcode <- factor(all_data$postcode, ordered=TRUE)
all_data$state <- factor(all_data$state, ordered=TRUE)
all_data$pc_immun <- factor(all_data$pc_immun, ordered=TRUE)
all_data$PHN_code <- factor(all_data$PHN_code, ordered=TRUE)
all_data$electorate <- factor(all_data$electorate, ordered=TRUE)
str(all_data)
View(all_data)
```

##Now let's build a more complete model using the two years to see if things can be confirmed
```{r}
#mixed model of both years
imm.model.mixed <- 'imm16 =~ pc_immun_class + IRSAD_SCORE + political_score + govt_funds_per_person + mean_tax_000s
  imm15 =~ pc_immun_class + IRSAD_SCORE + political_score + govt_funds_per_person + mean_tax_000s'
mixed.fit <- cfa(model = imm.model.mixed,
                  data = all_data)
```

```{r}
```
```{r}
```
```{r}
```


