---
title: "LavaanModellingOct15"
author: "Alex Brooks"
date: "10/15/2018"
output: html_document
---

```{r setup, include=FALSE}
library(lavaan)
library(semPlot)
options (scipen = 999)
library(dplyr)
library(caret)
library(corrplot)
library(glmnet)
library(gapminder)
library(openintro)
library(ggplot2)
library(Amelia)
library(missMDA)
```
 
 

## Let's just look at trying to make some single factor lavaan models work  

Start with 2016 only data

```{r 2016}
 #read in the data for 2016 only
Data <-read.csv('../cleaned_data/2016_data.csv')
str(Data)
```
```{r}
#Filter and clean data to get only the variables we really want

geo_data <- Data %>%
  select(postcode, state, year, age, pc_immun, caution, pc_immun_class, PHN_number, IEO_SCORE, IER_SCORE, IRSAD_SCORE, IRSD_SCORE, electorate, mean_tax_000s, political_score)
str(geo_data)
```
##Remove zeros from pc_immun_class
This seems an impossible task - ask Kirsty or Jarod tonight
```{r cleaning}
# Remove NA values
Data <- Data[complete.cases(Data), ]

#Need to make the factors ordered factors to work in Lavaan
Data$postcode = postcode_factor_ordered <- factor(Data$postcode, ordered = TRUE)
Data$state = state_factor_ordered <- factor(Data$state, ordered = TRUE)
Data$pc_immun = pc_immun_factor_ordered <- factor(Data$pc_immun, ordered = TRUE)
Data$PHN_code = PHN_code_factor_ordered <- factor(Data$PHN_code, ordered = TRUE)
Data$electorate = electorate_factor_ordered <- factor(Data$electorate, ordered = TRUE)
glimpse(Data)
```

```{r, fig.width=18, fig.height=12}
#correlation plot - this won't work, I think due to NAs
cplot <- geo_data %>%
  select_if(is.numeric)


M <- cor(cplot)
p.mat <- cor.mtest(M)
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
corrplot(M, 
         method = "color",
         order= "hclust",
         type="full",
         col=col(200),
         diag =F,
         title="Correlation of Numeric Variables",
         addCoef.col = "black",
         sig.level = 0.05,
         insig ="blank",
         mar=c(0,0,3,0))

```
 

#First 2016 model - not very good
```{r}
pc.model <- 'pc_immun_class =~ postcode + PHN_number + IRSAD_SCORE + electorate + state'
pc.fit <- cfa(model=pc.model,
                  data = Data)
summary(pc.fit, standardized = TRUE, fit.measures = TRUE)
#PHn number could be linearly regressed - does not make sense 

```
```{r}
lavInspect(pc.fit, what="estimates")
```

```{r}
semPaths(pc.fit,  what = "stand", rotation = 4)
```

You can also embed plots, for example:

```{r}
semCors(pc.fit)
```

#Try the same single factor model as first model usng 2015 data  
```{r 2015}
 #read in the data for 2015 only
#try this to stop data confusion
char_data <- read.csv('../cleaned_data/2015_data.csv', stringsAsFactors=F)
num_data <- data.frame(data.matrix(char_data))
numeric_columns <- sapply(num_data,function(x){mean(as.numeric(is.na(x)))<0.5})
final_data <- data.frame(num_data[,numeric_columns], char_data[,!numeric_columns])
str(final_data)
```
 

```{r}
#Our geo factor is measured by PHN, IRSAD score and pc immun class
geo_2015.model <- 'pc_immun_class =~ postcode + PHN_number + IRSAD_SCORE'
geo_2015.fit <- cfa(model = geo_2015.model,
                  data = final_data)
summary(geo_2015.fit, standardized = TRUE, fit.measures = TRUE)
  
```


```{r}
lavInspect(geo_2015.fit, what="estimates")
```
```{r}
semPaths(geo_2015.fit,  what = "stand", rotation = 4)
```

You can also embed plots, for example:

```{r}
semCors(geo_2015.fit)
```
```{r}
#modification indices (rmsea and srmr need to be below 0.10)
#look at loading and variances to check whether they are improbable or just don't relate to the latent variable, if loadings are over point 3 (variance will be large in comparison to the data)
#var(final_data)

```

#Not a great geo model
QUESTIONS:
How can I correctly set the weight of these to become properly measured manifest variables?

#Third model
```{r}
#Try a single factor model on the sentiment of trust

trust.model <- 'score =~ IRSAD_SCORE + political_score + mean_tax_000s + postcode'
trust.fit <- cfa(model=trust.model,
                  data = geo_data)
summary(trust.fit, standardized = TRUE, fit.measures = TRUE)
  
```
```{r}

semCors(trust.fit)
```

```{r}
semPaths(trust.fit,  what = "stand", rotation = 4)
```
```{r}
semCors(trust.fit, include="difference")
```
##This isn't really working
Let's see if we can replicate the geo modelling at SA3 level
##Fourth model
```{r}
#try this to stop data confusion
char_data <- read.csv('../raw_data/sa3_immune.csv', stringsAsFactors=F)
num_data <- data.frame(data.matrix(char_data))
numeric_columns <- sapply(num_data,function(x){mean(as.numeric(is.na(x)))<0.5})
final_data <- data.frame(num_data[,numeric_columns], char_data[,!numeric_columns])
str(final_data)
```


```{r}
 
#Try a single factor model
#Our geo factor is measured by PHN, IRSAD score and pc immun class
sa3.model <- 'sa3 =~ SA3_code + year_start + age + num_not_fully_immunised + num_fully_immunised + percent_fully_immunised'
sa3.fit <- cfa(model=sa3.model,
                  data = final_data)
summary(sa3.fit, standardized = TRUE, fit.measures = TRUE)
  
```

```{r}
#plot it
lavInspect(sa3.fit, what="estimates")
```
```{r}
semPaths(sa3.fit,  what = "stand", rotation = 4)
```
```{r}
semCors(sa3.fit, include="difference")
```
